### 2015-02-11 Projects/DelCo_2012/Report/DelCo_2012.tex
DelCo 2012: Minor text updates.

>>>>

The National Institute of Standards and Technology, with the support of the Fire Protection Research
Foundation, the U.S. Department of Homeland Security, and the U.S. Fire Administration conducted a
series of fire experiments to examine the impact of wind on fire spread through a multi-room structure
and examine the capabilities of wind-control devices (WCD) and externally applied water to mitigate the
hazard. The measurements used to examine the impact of the WCDs and the external water application
tactics were heat release rate, temperature, heat flux, and gas velocity inside the structure. Oxygen,
carbon dioxide, carbon monoxide, total hydrocarbons and differential pressures were also measured.
Each of the experiments was recorded with video and thermal imaging cameras. Some of these
measurements are not practical or affordable to make in an acquired structure, hence the need to build a
structure and conduct the experiments within the confines of the NIST Large Fire Facility. These 354
experiments also provided visual documentation of fire phenomena that are not typically observable on
the fire ground.
A limited series of heat release rate experiments were conducted to characterize the fuel load packages
used in wind driven structure experiments. Both the bedroom and the living room contained a fuel load
composed of furnishings with an average peak heat release rate of 7.8 MW with a total heat release of at
least 1700 MJ, not accounting for any of the wooden furniture or interior finish materials.
The experiments were designed to expose a public corridor area to a wind driven, post-flashover
apartment fire. The door from the apartment to the corridor was open for each of the experiments. The
conditions in the corridor were of critical importance because that is the portion of the building that
firefighters would use to approach the fire apartment or that occupants from an adjoining apartment
would use to exit the building.
The fires were ignited in the bedroom of the apartment. Prior to the failure or venting of the bedroom
window, which was on the upwind side of the experimental apartment, the heat release rate from the fire
was on the order of 1 MW. Prior to implementing either of the mitigating tactics, the heat release rates
from the post-flashover structure fire were typically between 15 MW and 20 MW. When the door from
the apartment to the corridor was open, temperatures in the corridor area near the open doorway, 1.52 m
(5.00 ft) below the ceiling, were in excess of 600 °C (1112 °F) for each of the experiments. The heat
fluxes measured in the same location, during the same experiments, were in excess of 70 kW/m². These
extreme thermal conditions are not tenable, even for a firefighter in full protective gear. These
conditions were attained within 30 s of the window failure.
Experiment 1 was conducted without any external wind. This experiment provided valuable baseline
data and demonstrated several important points relevant to fire fighting:
Smoke is Fuel. A ventilation limited (fuel rich) condition had developed prior to the failure of
the window. Oxygen depleted combustion products, containing carbon dioxide, carbon monoxide and
unburned hydrocarbons filled the rooms of the structure. Once the window failed, the fresh air provided
the oxygen needed to sustain the transition through flashover, which caused a significant increase in heat
release rate.
Venting does not always equal cooling. In this experiment, post ventilation temperatures and
heat fluxes all increased, due to the ventilation induced flashover.
Fire induced flows. Velocities within the structure exceeded 5 m/s (11 mph), just due to the fire
growth and the flow path that was set-up between the window opening and the corridor vent.
Avoid the flow path. The directional nature of the fire gas flow was demonstrated with thermal
conditions, both temperature and heat flux, which were twice as high in the “flow” portion of the
corridor as opposed to the “static” portion of the corridor in Experiment 1. Thermal conditions in the
flow path were not consistent with firefighter survival.
Experiments 2 through 8 all used a mechanically generated wind, ranging from 3 m/s to 9 m/s (7 mph to
20 mph). 355
The fuel load in the structure was the same for all of the experiments. Each of these experiments
demonstrated a rapid transition to untenable conditions in the corridor, even for a firefighter in full PPE,
after the window failed.
Experiments 2 through 5 focused on the impact of WCDS. In these experiments, the WCDs reduced
the temperatures in the corridor outside the doorway by more than 50 % within 60 s of deployment. The
heat fluxes were reduced by at least 70 % during this same time period. The WCDs also completely
mitigated any gas velocity due to the external wind.
Experiments 6 through 8 focused on the impact of externally applied water. In these experiments, the
externally applied water streams were implemented in three different ways; a fog stream across the face
of the window opening, a fog stream into the window opening, and a solid water stream into the window
opening. The fog stream across the window was not effective at reducing the thermal conditions in the
corridor. The fog stream in the window decreased the corridor temperature by at least 20 % and the
corresponding heat flux measured by at least 30 %. The solid stream experiments resulted in corridor
temperature and heat flux reductions of at least 40 % within 60 s of application. None of the water
applications reduced the gas velocities in the structure. In some cases, the gas velocity increased during
water application, due to momentum imparted from the water.
These experiments demonstrated the “extreme” thermal conditions that can be generated by a “simple
room and contents” fire and how these conditions can be extended along a flow path within a structure
when wind and an open vent are present. Two potential tactics which could be implemented from either
the floor above the fire in the case of a WCD, or from the floor below the fire in the case of the external
water application were demonstrated to be effective in reducing the thermal hazard in the corridor.
However, these experimental results also indicate that the post deployment thermal conditions for any
single tactic were still of a level which could pose a hazard to firefighters in full PPE.
The experiments also provided potential guidance for firefighters as a part of a fire size up and approach
to the room of fire origin: note wind conditions in the area of the fire, look for “pulsing flames”,
examine smoke conditions around closed doors in the potential flow path, and maintain control of doors
in the flow path.
Further research in actual buildings is required to fully understand the ability of firefighters to
implement these tactics, to examine the thermal conditions throughout the structure such as in stairways,
and to examine the interaction of these tactics with building ventilation strategies both natural and with
positive pressure ventilation.
If the demonstrated technologies continue to prove effective in the field trials and pilot programs, the
next step may be to examine the need for standards and standardized test methods to define a minimum
level of acceptable performance of these devices.
